---
title: "hw04: Tidy data and joins"
output: html_document
---

Load the required packages:
```{r}
library(gapminder)
library(tidyverse)
```


# Ex1. Univariate data reshaping
## Option 1 - Exploring life expectancy for Australia and Japan
### A tibble with one row per year, and columns for life expectancy for two or more countries.
<!-- Need to select and filter before widening the dataframe --> 
```{r}
(Ex1<- gapminder %>% 
  select(country, lifeExp, year) %>%
  filter(country == "Australia" | country == "Japan") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp))

```


### A scatter plot of the life expectancy of Japan against Australia
```{r}
ggplot(Ex1, aes(Australia, Japan))+
  geom_point()+
  ggtitle("Life expectancy for Japan and Australia")+
  theme_bw()
```
#### *Figure 1. Life expectancy of Japan against Australia.* 


### Re-lengthen the data
```{r}
Ex1 %>%
  pivot_longer(cols = c(Australia, Japan),
               names_to = "country",
               values_to = "lifeExp")
```


# Ex2. Multivariate data re-shaping
## Option 1
### A tibble for the life expectancy and GDP per capita for Australia and Japan
```{r}
(Ex2<- gapminder %>% 
  select(country, lifeExp, year, gdpPercap) %>%
   filter(country == "Australia" | country == "Japan") %>%
  pivot_wider(id_cols = year,
              names_from = c(country),
              names_sep = "_",
              values_from = c("lifeExp", "gdpPercap")))

```


### Re-lengthen the data
```{r}
Ex2 %>%
  pivot_longer(cols =  -year,
               names_to = c(".value", "country"),
               names_sep = "_")

```


# Ex3. Table joins
First read in the data from github
```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## 3.1 For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.
First I created an object that is the email list in long format, with only one observation (name) per e-mail. The object 'email.guest' is then created by joining the guest list with the email list by name.
```{r}
(email.long <- email %>% 
  separate_rows(guest, sep = ", ") %>%
   rename(name = guest))

(email.guest <- guest %>%
  left_join(email.long, by = "name"))
```


## 3.2 Who do we have emails for, yet are not on the guestlist?
```{r}
email.long %>%
  anti_join(guest)
```


## 3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).
This is a list that is a full guestlist of everyone there are emails for. I joined the email.long dataframe with the guest dataframe so that everyone we have emails is included. 
<!--Use full join for complete list (EVERYONE!)-->
```{r}
(full.list <- guest %>%
  full_join(email.long))
```

