---
title: "Exploring the gapminder dataset"
output: html_document
---
```{r, echo=FALSE}
library(gapminder)
library(dplyr)
library(ggplot2)
```
## _Basic dplyr_
### 1.1 Canada, Belgium and Bangladesh in the 1970s
```{r, echo=FALSE}
gapminder %>%
  filter(year >= "1970" & year < "1980") %>%
  filter(country == "Canada" | country == "Belgium" | country == "Bangladesh")
```

### 1.2 Previous figure with only country and gdpPercap
```{r, echo=FALSE}
gapminder %>%
  filter(year >= "1970" & year < "1980") %>%
  filter(country == "Canada" | country == "Belgium" | country == "Bangladesh")%>%
  select(country, gdpPercap)
```

### 1.3 All entries with a drop in life expectancy
<!-- Lag allows you to calculate running differences between observations -->
```{r, echo=FALSE}
gapminder %>%
  arrange(year) %>%
  mutate(life.inc = lifeExp - lag(lifeExp)) %>%
  filter(life.inc <0)
```

### 1.4 Max GDP experienced per country
```{r, echo=FALSE}
gapminder %>%
  group_by(country) %>%
  summarise(max.gdp = max(gdpPercap))
```

### 1.5 Scatter plot of Canada's life expectancy vs. GDP per capita
```{r, echo=FALSE}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(., aes(gdpPercap, lifeExp))+
           geom_point()+
  scale_x_log10("GDP per capita", labels = scales::dollar_format())+
  ylab("Life expectancy (years)")+
  theme_classic()
```

##  _2 Exploring year and life expectancy_ 
### Simple summary statistics of continent (categorical variable) and life expectancy (quantitative variable)
```{r, echo=FALSE}
gapminder %>%
  select(continent, lifeExp)%>%
  summary()
```

### The number of countries per continent and, mean, max. and min. life expectancy for each continent 
```{r, echo=FALSE}
gapminder %>%
  select(lifeExp,continent) %>%
  group_by(continent) %>%
  summarise(n.of.countries = n(),
            yearly.maxLE = max(lifeExp),
            yearly.meanLE = mean(lifeExp),
            yearly.minLE = min(lifeExp))
```

### Line graph depicting the increase in life expectancy in each continent over time. Grey areas are standard error
```{r, echo=FALSE}
gapminder %>%
  select(year, lifeExp, continent) %>%
  ggplot(., aes(year, lifeExp, group = continent))+
           geom_smooth(method = "loess", aes(color = continent))+
  labs(title = "Growth in life expectancy over time by continent", x = "Year", y = "Life expectancy (years)")+
  theme_bw()
      
```




## _3 Explore various plot types_
### Scatter plot of population and GDP per capita for Australia, China, Japan and India
```{r, echo=FALSE}
gapminder %>%
  filter(country =="Australia" | country == "China"  | country == "Canada" | country == "India" ) %>%
  ggplot(., aes(gdpPercap, pop, group=country))+
  scale_x_continuous("GDP per capita", labels = scales::dollar_format())+
  ylab("Population")+
  geom_point(aes(color=country))+
  geom_line(aes(color=country))+
  theme_classic()
```

## Growth of GDP per capita over time 
```{r, echo=FALSE}
gapminder %>%
  filter(country =="Australia" | country == "China"  | country == "Canada" | country == "India" ) %>%
  ggplot(., aes(year, gdpPercap))+
  scale_y_log10("GDP per capita", labels = scales::dollar_format())+
  geom_line(aes(color=country))+
  facet_wrap(~country)+
  xlab("Year")+
  ggtitle("Growth of GDP per capita over time")+
  theme_classic()
```





