---
title: "Exploring gapminder with dplyr"
output: html_document
---

```{r}
library(gapminder)
library(dplyr)
<<<<<<< HEAD
=======
<<<<<<< HEAD
>>>>>>> a0c7e74316b082a9e3cc1d597360c21ec0115c56
library(ggplot2)
```

## EX 1.1 Canada, Belgium and Bangladesh in the 1970s
=======
```

## 1.1 Canada, Belgium and Bangladesh in the 1970s
<<<<<<< HEAD
=======
>>>>>>> a66da7f32ddbc6771cdd1d2d4e7996fb41aea568
>>>>>>> a0c7e74316b082a9e3cc1d597360c21ec0115c56
```{r}
gapminder %>%
  filter(year >= "1970" & year < "1980") %>%
  filter(country == "Canada" | country == "Belgium" | country == "Bangladesh")
```

## 1.2 Above with only country and gdpPercap
```{r}
gapminder %>%
  filter(year >= "1970" & year < "1980") %>%
  filter(country == "Canada" | country == "Belgium" | country == "Bangladesh")%>%
  select(country, gdpPercap)

```

## 1.3 All entries with a drop in life expectancy
```{r}
gapminder %>%
<<<<<<< HEAD
  mutate(life.inc = c(NA, diff(lifeExp, lag = 1))) %>%
  filter(life.inc < 0)
=======
<<<<<<< HEAD
  mutate(life.inc = c(NA, diff(lifeExp, lag = 1))) %>%
  filter(life.inc < 0)
=======
>>>>>>> a66da7f32ddbc6771cdd1d2d4e7996fb41aea568

>>>>>>> a0c7e74316b082a9e3cc1d597360c21ec0115c56
```

## 1.4 Max GDP experienced per country
```{r}
gapminder %>%
  group_by(country) %>%
<<<<<<< HEAD
=======
<<<<<<< HEAD
>>>>>>> a0c7e74316b082a9e3cc1d597360c21ec0115c56
  summarise(max.gdp = max(gdpPercap))
  
```

## 1.5 Scatter plot of Canada's life expectancy vs. GDP per capita
```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(., aes(gdpPercap, lifeExp))+
           geom_point()+
  scale_x_log10("GDP per capita", labels = scales::dollar_format())+
  ylab("Life expectancy")+
  theme_bw()
```

## EX 2 Exploring country and population
```{r}
gapminder %>%
  select(country, pop) %>%
  summarise()
```

#EX 3 Plotting
```{r}
ggplot(trees, aes(Girth, Volume))+
  geom_point()+
  theme_classic()
```

<<<<<<< HEAD


=======
=======
  mutate(maxGDP.country = max(gdpPercap))

gapminder %>%
  group_by(country) %>%
  summarise(max.gdp = max(gdpPercap))
  
```

## 1.5 
>>>>>>> a66da7f32ddbc6771cdd1d2d4e7996fb41aea568
>>>>>>> a0c7e74316b082a9e3cc1d597360c21ec0115c56
